---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

**NOTE: This is currently in development. **

# BiodiverseR    

<!-- badges: start -->
[![R-CMD-check](https://github.com/shawnlaffan/Biodiverse-R/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/shawnlaffan/Biodiverse-R/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Provides an R interface to the analyses available in Biodiverse. Biodiverse 
is a tool for the spatial analysis of diversity using indices based on taxonomic, 
phylogenetic, trait and matrix-based (e.g. genetic distance) relationships, 
as well as related environmental and temporal variations. More information is 
available at its [Github page](https://github.com/shawnlaffan/biodiverse).

## Installation

This is a two step process.  First, install the Biodiverse engine.  Second, install the R package.  

1. Install the Biodiverse engine. 

This currently requires a working perl interpreter in your path.
(Future versions will provide self contained executables).

On Windows a perl interpreter can be obtained through the [Strawberry perl project](https://strawberryperl.com/releases.html).
This will be downloaded automatically when using the commands below.  

Most unix-derived systems provide a perl interpreter but it is best to avoid this and install 
[perlbrew](https://perlbrew.pl/) so you have a separate installation.  
When you install perlbrew be sure to also install the cpanm utility (see
perlbrew site for details).  

Use perlbrew to install a recent version of perl.

You also need to have git installed on your system and in the path.


2. Install the R code


You can install the R code like so:

``` r
library("devtools")
devtools::install_github("shawnlaffan/BiodiverseR")
```

However, it is currently best to work within the git repo.  
Set your working directory to be the top of the git repo and then run this:

```r
library("devtools")
devtools::load_all()
```

To install the perl dependencies, run these commands.  
The first one does nothing on Windows but there is no harm in running it.  

```r
init_perlbrewr()
install_perl_deps()
```

Note that the above will take a while if you do not already have the GDAL development
package installed on your system.  This is because it will compile its own version if
it is unable to find one on the system (but maybe this is not such a bad thing as then
it will be isolated from system changes).



## Quick demo

Check that the Biodiverse service can be accessed.  The analytical functions
call this internally so this is just a check that the server can be started.

```{r}
# It is critical that this be set to wherever you have downloaded the package 
#  as otherwise the system will not find the server code.  
#  It is an ugly and temporary hack and will not be needed in the future.
#  This version assumes you are at the top level of the BiodiverseR repository.  
Sys.setenv("Biodiverse_basepath" = getwd())

#  library(BiodiverseR)
devtools::load_all()  #  for during development 
cs = start_server()
cs$server_object$is_alive()

#  cleanup
rm(cs)
gc()  #  server is not deleted until garbage collected
```
