---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

**NOTE: This is a package template, which will be used to provide an interface to the Biodiverse tool. **

# BiodiverseR    

<!-- badges: start -->
[![R-CMD-check](https://github.com/shawnlaffan/Biodiverse-R/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/shawnlaffan/Biodiverse-R/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Provides and interface to functionalities available in the Biodiverse. Biodiverse 
is a tool for the spatial analysis of diversity using indices based on taxonomic, 
phylogenetic, trait and matrix-based (e.g. genetic distance) relationships, 
as well as related environmental and temporal variations. More information is 
available at its [Github page](https://github.com/shawnlaffan/biodiverse).

## Installation

This currently requires a working perl interpreter in your path.
(Future versions will provide self contained executables).

On Windows perl interpreter  can be obtained through the [Strawberry perl project](https://strawberryperl.com/releases.html).
Most unix-derived systems provide a perl interpreter but it is best to avoid this and use
a system like [perlbrew](https://perlbrew.pl/).  Be sure to also install the cpanm utility (see
perlbrew site for details).  

Once you have a perl installed and in your path you can install the perl dependencies using
cpanm at the command line.  Make sure to update this code to use the correct path separator
on Windows.

``` perl
cpanm https://github.com/shawnlaffan/biodiverse.git
cd inst/perl
cpanm --installdeps .
```

Note that the above will take a while if you do not already have the GDAL development
package installed on your system.  This is because it will compile its own version if
it is unable to find one on the system (and maybe this is not such a bad thing as then
it will be isolated from system changes).

You can install the R code like so:

``` r
devtools::install_github("shawnlaffan/BiodiverseR")
```
  
## Quick demo

Checking that the Biodiverse service can be accessed:

```{r}
#  An ugly and temporary hack so we can find the relevant perl code.
#  Set this to wherever you have downloaded the package to.  
Sys.setenv("Biodiverse_basepath" = getwd())

#  library(BiodiverseR)
devtools::load_all()  #  for during development 
cs = start_server()
cs$process_object$is_alive()

#  cleanup
rm(cs)
gc()  #  server is not deleted until garbage collected
```
